---
title: "SuperCellBM (SuperCell benchmarking)"
csl: elsevier-harvard.csl
output:
  md_document:
    variant: markdown_github
---

An example how to build multiple super-cell-like objects, including 'exact' (Super-cells obtained with the exact coarse-gaining), 'approx' (Super-cells obtained with the aaproximate coarse-graining), 'metacell(.*)' (Metacell build on the same genes as super-cells -- 'metacell_SC_like'; and Metacell build in a default set of genes -- 'metacell_default') for a set of graining levels and random seeds.

```{r libraries, message=FALSE, warning=FALSE}
# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("SingleCellExperiment")
# 
# if (!requireNamespace("remotes")) install.packages("remotes")
# remotes::install_github("GfellerLab/SuperCell")
# remotes::install_github("mariiabilous/SuperCellBM")

library(SingleCellExperiment)
library(SuperCell)
library(SuperCellBM)
```


## Load some default parameters
Such as `.gamma.seq`for the set of fraining levels, `.seed.seq` for the set of random seeds, `adata.folder` and `fig.folder` for the folders where to write data and plots
```{r config files}
source("./examples/config/Tian_config.R")
```

### Flags
whether to compute super-cell (`ToComputeSC`) or whether to compute super-cell gene expression (`ToComputeSC_GE`) or load saved files. Make sure, this file exists :) 
```{r flags}
ToComputeSC <- T
ToComputeSC_GE <- T
```

##Load `cell_lines` data from \link[Tian et al., 2019](https://pubmed.ncbi.nlm.nih.gov/31133762/).
```{r load data }
RData.file.path <- file.path(data.folder, 'cell_lines_git.RData')

if(!file.exists(RData.file.path)){
  if(!dir.exists(data.folder)) dir.create(data.folder, recursive = T)
  download.file('https://github.com/LuyiTian/sc_mixology/blob/master/data/sincell_with_class_5cl.RData?raw=true', 
                RData.file.path)
}

load(RData.file.path)

# keep used dataset 
cell_lines_SCE <- sce_sc_10x_5cl_qc

#remove not-used datasets
rm(sc_Celseq2_5cl_p1, sc_Celseq2_5cl_p2, sc_Celseq2_5cl_p3, sce_sc_10x_5cl_qc)
```

Get and set the main variables, such as single-cell gene expression (`sc.GE`), single-cell counts (`sc.counts`), 
number of single cells (`N.c`) and total number of genes (`N.g`).
Set matrix column names to cellIDs (`cell.ids`) and row names to gene names (`gene.names`)
```{r get main values from the dataset}
cell.ids     <- cell_lines_SCE@colData@rownames
N.c          <- cell_lines_SCE@colData@nrows 


gene.names   <- cell_lines_SCE@int_elementMetadata$external_gene_name
N.g          <- length(gene.names)

sc.GE           <- cell_lines_SCE@assays$data$logcounts
colnames(sc.GE) <- cell.ids
rownames(sc.GE) <- gene.names

sc.counts   <- cell_lines_SCE@assays$data$counts
colnames(sc.counts) <- cell.ids
rownames(sc.counts) <- gene.names

## this is not needed at this point, but will be used later
GT.cell.type <- cell_lines_SCE@colData$cell_line_demuxlet
names(GT.cell.type) <- cell.ids
N.clusters         <- length(unique(GT.cell.type))

GT.cell.type.names          <- names(table(GT.cell.type))
GT.cell.type.2.num          <- 1:length(unique(GT.cell.type))
names(GT.cell.type.2.num)   <- GT.cell.type.names
GT.cell.type.num            <- GT.cell.type.2.num[GT.cell.type]
names(GT.cell.type.num)     <- names(GT.cell.type)

## uncomment this when needed 
#.pal.GT <- .color.tsne.Tian  ## to Global Config
#scales::show_col(.pal.GT)


#mito.genes <-  grep(pattern = "^MT", x = gene.names, value = TRUE)
#ribo.genes <-  grep(pattern = "^RP[LS]", x = gene.names, value = TRUE)
#mito.ribo.genes <- c(mito.genes, ribo.genes)
#length(mito.ribo.genes)

#gene.meta <- data.frame(name = gene.names, inNcells = rowSums(sc.GE>0), mean.expr = rowMeans(sc.GE), sd = rowSds(sc.GE))
#head(gene.meta)

```

## Compute Super-cell structure
for the Exact, Aprox (Super-cells obtained with the exact or approximate coarse-graining), Subsampling or Random (random grouping of cells into super-cells)

```{r compute super-cell structure}
  
SC.list <- compute_supercells(
  sc.GE,
  ToComputeSC = ToComputeSC,
  data.folder = data.folder,
  filename = 'initial',
  gamma.seq = .gamma.seq,
  n.var.genes = .N.var.genes,
  k.knn = .k.knn,
  n.pc = .N.comp,
  approx.N = 1000,
  fast.pca = TRUE,
  genes.use = .genes.use, 
  genes.exclude = .genes.omit,
  seed.seq = .seed.seq
  )

cat(paste("Super-cell computed for:", paste(names(SC.list), collapse = ", "), 
          "\nat graining levels:", paste(names(SC.list[['approx']]), collapse = ", "),
          "\nfor seeds:", paste(names(SC.list[['approx']][[1]]), collapse = ", "), "\n"))
```

## Compute metacells in two settings:
- 'metacell_default' - when metacell is computed with the default parameters (from the tutorial), using gene set filtered by MC
- 'metacell_SC_like' - when metacell is computed with the same default parameters, but at the same set of genes as Super-cells 

```{r computr metacells}
SC.mc <- compute_supercells_metacells(
  sc.counts = sc.counts, 
  gamma.seq = .gamma.seq,
  SC.list = SC.list,
  proj.name = proj.name,
  ToComputeSC = ToComputeSC, 
  mc.k.knn = 100,
  T_vm_def = 0.08,
  MC.folder = "MC", 
  MC_gene_settings = c('metacell_default', 'metacell_SC_like') # do not change
)

additional_gamma_seq <- get_actual_gammas_metacell(SC.mc)

cat(paste("Metacells were computed in", length(names(SC.mc)), "settings:", paste(names(SC.mc), collapse = ", "), "\n",
          "For Gammas:", paste(names(SC.mc[[1]]), collapse = ", "), "\n",
          "But actual gammas are:", paste(additional_gamma_seq, collapse = ", "), "\n"
))


# manually expand MC because later we will have 2 diferetn setting for MC profile: fp - footpring of MC, av - averaged 
SC.mc.fp <- SC.mc
names(SC.mc.fp) <- sapply(names(SC.mc), FUN = function(x){paste0(x, '_fp')})

SC.mc.av <- SC.mc
names(SC.mc.av) <- sapply(names(SC.mc), FUN = function(x){paste0(x, '_av')})

SC.mc.expanded <- c(SC.mc.fp, SC.mc.av)

names(SC.mc.expanded)
rm(SC.mc.fp, SC.mc.av, SC.mc)
```

## Compute super-cells (Exact, Approx, Subsampling and Random) at the addidional grainig levels obtained with Metacell.
So that we can dirrectly compare the results of Super-cells and Metacells at the same graining levels.

```{r compute SC at additionla gammas}
SC.list <- compute_supercells_additional_gammas(
  SC.list,
  additional_gamma_seq = additional_gamma_seq,
  ToComputeSC = TRUE,
  data.folder = data.folder,
  filename = 'additional_gammas',
  approx.N = 1000,
  fast.pca = TRUE
)

print(paste("Super-cells of methods:", paste(names(SC.list), collapse = ", "), 
      "were computed at aggitional graining levels:", paste(additional_gamma_seq, collapse = ", "), "
      and added to SC.list"
      ))
```

### Concatenate Metacells to the list of Super-cells

```{r concat super-cell and metacells}

SC.list <- c(SC.list, SC.mc.expanded)
names(SC.list)
rm(SC.mc.expanded)

```

## Compute GE for Super-cell data 
GE profile for the super-cell data is computede:
- for super-cells (Exact, Approx) by averaging gene expression within super-cells,
- for Random, also averaging gene expression within super-cells 
- for the Subsampling, sc.GE matrix is just subsampled,
- for Metacells, gene expression is computed in 2 ways: 
1) the same as super-cells (averaging gene expression within Metacells) -> `metacell_(.*)_av`
2) using the default output of Metacell maned footprint -> `metacell_(.*)_fp`


```{r compute GE for all super cell like objects}

SC.GE.list <- compute_supercells_GE(
  sc.GE = sc.GE, 
  SC.list = SC.list,
  ToComputeSC_GE = ToComputeSC_GE, 
  data.folder = data.folder,
  filename = 'all'
)

cat(paste("Gene expression profile computed for:", paste(names(SC.GE.list), collapse = ", "), 
    "\nat graining levels:", paste(sort(as.numeric(names(SC.GE.list[['Approx']]))), collapse = ", "),
    "\nfor seeds:", paste(names(SC.GE.list[['Approx']][[1]]), collapse = ", ")))
```


